---
title: "SVM plots"
author: "Liza Brusman"
date: "2024-10-29"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r, message = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggridges)
library(forcats)
library(ggpubr)
library(glmmTMB)
library(DHARMa)
library(gridExtra)
library(emmeans)
library(Hmisc)
library(corrplot)
source("SVM_plot_fxns.R")
```

metadata
```{r}
metadata <- read.csv("../../docs/seq_beh_metadata.csv")
metadata$Group <- paste(metadata$sex, metadata$SS_OS, sep = "_")
metadata <- metadata %>% filter(!animal %in% c("4918", "4967"))
```


module data
```{r}
all_cells <- read.csv("../hotspot/docs/ani_mod_scores_allcells_lognorm_counts.csv")
modules <- colnames(all_cells)
modules <- modules[modules != "animal"]

data <- merge(all_cells, metadata, on = "animal")
```


for SVMs that include self:
```{r}
with_self <- read.csv("output/ani_pairwise_classifications_withself.csv")
```

for SVMs that do not include self
```{r}
exclude_self <- read.csv("output/all_svm_combined_excludeself_200cells.csv") %>% filter(animal == Var1)
```

put clusters in right order and assign colors
```{r}
f.levels <- c('Drd1Pdyn', 'Drd1PdynOprm1', 'Drd1Penk', 'Drd2Penk', 'Drd2NoPenk', 
                                      'GABAergicNeurons', 'Dlx2ImmatureNeurons', 'SstNpyInterneurons', 
                                      'PvalbInterneurons', 'CholinergicInterneurons', 'MatureOligos', 
                                      'ImmatureOligos', 'Astrocytes', 'Microglia')

clust_cols <- c("Drd1Pdyn" = "#304880", "Drd1PdynOprm1" = "#4898D0", "Drd1Penk" = "#88C0F0", "Drd2Penk" = "#9060A0", "Drd2NoPenk" = "#C078C0", "GABAergicNeurons" = "#389078", "Dlx2ImmatureNeurons" = "#70A830", "SstNpyInterneurons" = "#98D048", "PvalbInterneurons" = "#60D0A0", "CholinergicInterneurons" = "#80E8C0", "MatureOligos" = "#783028", "ImmatureOligos" = "#B82820", "Astrocytes" = "#D04058", "Microglia" = "#F87878")
```

process df into format for ridgeline plots
```{r}
with_self2 <- process_df_withself(with_self)
```


plot ridgeline plot
```{r}
plot_ridgeline(with_self2, save = FALSE)
```
STAT TESTS - GLM
for with self - self vs. partner vs. other
```{r}
with_self_stats <- GLM_SVM(with_self2)

# setwd("output/")
# write.csv(with_self_stats, "glm_svm_with_self_groups_ordbeta.csv")
```

widen dataframe for other analyses later
```{r}
with_self_wide <- widen_df_withself(with_self2)
```

process df for excluding self
```{r}
exclude_self2 <- process_df_excludeself(exclude_self)
```
plot exclude self ridgeline
```{r}
plot_ridgeline(exclude_self2, save = FALSE)
```

GLM stats: exclude self
```{r}
exclude_self_stats <- GLM_SVM(exclude_self2)

# setwd("output/")
# write.csv(exclude_self_stats, "svm_exclude_self_glm_groups.csv")

```

exclude self widened df
```{r}
exclude_self_wide <- widen_df_excludeself(exclude_self2)
```

plot ridgelines by pairing type - with self
```{r}
pair_types <- unique(data$pair_type)
for (type in pair_types) {
  print(type)
  plot_ridgeline(with_self2, grouping = "pair_type", save = FALSE)

}
```
ridgeline by pair type without self
```{r}
pair_types <- unique(data$pair_type)
for (type in pair_types) {
  print(type)
  plot_ridgeline(exclude_self2, grouping = "pair_type", save = FALSE)
  
}
```
ridgeline by cohort with self
```{r}
cohs <- unique(data$beh_cohort)
for (type in cohs) {
  print(type)
  plot_ridgeline(with_self2, grouping = "beh_cohort", save = FALSE)
  
}
```
ridgelines by cohort exclude self
```{r}
cohs <- unique(data$beh_cohort)
for (type in cohs) {
  print(type)
  plot_ridgeline(exclude_self2, grouping = "beh_cohort", save = FALSE)
  
}

```

directly compare pairing types. each point in these plots is the percentage of classification for one animal to one other animal
```{r}
exclude_self2$group_type <- paste(exclude_self2$pair_type, exclude_self2$group, sep = "_")

plot_dotcloud(exclude_self2)

```

for stats - looking at partner/other*pair_type
```{r}
exclude_self_pairtype_stats <- GLM_SVM_pairtype(exclude_self2)

# setwd("output/")
# write.csv(exclude_self_pairtype_stats, "svm_excludeself_glm_posthocs_by_pairtype.csv")

```


correlation between partners in partner % classification
```{r}
plot_SVM_corrs(exclude_self_wide, save = FALSE)
```

get out some summary stats for exclude self data
```{r}
summ_stats <- exclude_self_wide %>% group_by(Cluster) %>% summarise(mean_partner_pct = mean(partner.pct), avg_cells_tested = mean(partner+other))

ggplot(summ_stats, aes(x = avg_cells_tested, y = mean_partner_pct, color = Cluster)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_color_manual(values = clust_cols)

#here i correlated the mean partner % vs. the number of cells tested
rcorr(summ_stats$mean_partner_pct, summ_stats$avg_cells_tested)
```

get some summary stats for with self data
```{r}
summ_stats <- with_self_wide %>% group_by(Cluster) %>% summarise(mean_self_pct = mean(self.pct), avg_cells_tested = mean(self+partner+other))

#here i correlated the self classifications to the number of cells tested. this is highly correlated
cor_test <- rcorr(summ_stats$mean_self_pct, summ_stats$avg_cells_tested, type = "spearman")
print(cor_test)

ggplot(summ_stats, aes(x = avg_cells_tested, y = mean_self_pct, color = Cluster)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_color_manual(values = clust_cols)
```


self vs. chance - one-sample t-test vs. 1/38 (correct classification by random)
```{r}
vs_chance_df <- data.frame()

for (clust in f.levels) {
  clust_df <- with_self2 %>% filter(Cluster == clust) %>% filter(group == "self")
  
  onesamp <- t.test(clust_df$norm_classes, mu = 1/38)
  # print(oneway)
  
  stats_df <- data.frame(Cluster = clust,
                        T_val = onesamp$statistic[1],
                        p.value = onesamp$p.value,
                        test = "onesamp t-test")
  
  vs_chance_df <- rbind(vs_chance_df, stats_df)
  
}
```

partner vs. chance for exclude self
```{r}
vs_chance_df_exclude <- data.frame()

for (clust in f.levels) {
  clust_df <- exclude_self2 %>% filter(Cluster == clust) %>% filter(group == "partner")
  
  onesamp <- t.test(clust_df$norm_classes, mu = 1/37)
  # print(oneway)
  
  stats_df <- data.frame(Cluster = clust,
                        T_val = onesamp$statistic[1],
                        p.value = onesamp$p.value,
                        test = "onesamp t-test")
  
  vs_chance_df_exclude <- rbind(vs_chance_df_exclude, stats_df)
  
}
```


look at classifications for same sex, same cohort, same pairing type, relatedness, etc.
basically - what other factors are at play other than partner ID?
```{r}
exclude_self_newclass <- exclude_self2
exclude_self_newclass <- exclude_self_newclass %>% merge(metadata, on = "animal")

#basically just creating named lists to match to predicted animal
sex_vec <- metadata$sex
names(sex_vec) <- metadata$animal

exclude_self_newclass$y_pred <- as.character(exclude_self_newclass$y_pred)
exclude_self_newclass$same_opp_sex <- if_else(exclude_self_newclass$sex == sex_vec[exclude_self_newclass$y_pred], "same", "different")

coh_vec <- metadata$beh_cohort
names(coh_vec) <- metadata$animal
exclude_self_newclass$same_diff_coh <- if_else(exclude_self_newclass$beh_cohort == coh_vec[exclude_self_newclass$y_pred], "same", "different")

type_vec <- metadata$pair_type
names(type_vec) <- metadata$animal
exclude_self_newclass$same_diff_type <- if_else(exclude_self_newclass$pair_type == type_vec[exclude_self_newclass$y_pred], "same", "different")

group_vec <- metadata$Group
names(group_vec) <- metadata$animal
exclude_self_newclass$same_diff_group <- if_else(exclude_self_newclass$Group == group_vec[exclude_self_newclass$y_pred], "same", "different")

parent_vec <- metadata$parents
names(parent_vec) <- metadata$animal
exclude_self_newclass$same_diff_parents <- if_else(exclude_self_newclass$parents == parent_vec[exclude_self_newclass$y_pred], "same", "different")
```

only look at "other" classifications
```{r}
exclude_self_newclass_other <- exclude_self_newclass %>% filter(group == "other")
norm_to_other_class <- exclude_self_newclass_other %>% group_by(Cluster, animal) %>% summarise(sum_other_classes = sum(Freq))
same_sex_norm <- exclude_self_newclass_other %>% group_by(Cluster, animal, same_opp_sex) %>% summarise(classes = sum(Freq))
same_sex_norm$metric <- "sex"
same_sex_norm <- same_sex_norm %>% rename("same_diff" = "same_opp_sex")
same_sex_norm <- same_sex_norm %>% merge(norm_to_other_class, on = "animal")

same_coh_norm <- exclude_self_newclass_other %>% group_by(Cluster, animal, same_diff_coh) %>% summarise(classes = sum(Freq))
same_coh_norm$metric <- "cohort"
same_coh_norm <- same_coh_norm %>% rename("same_diff" = "same_diff_coh")
same_coh_norm <- same_coh_norm %>% merge(norm_to_other_class, on = "animal")

same_type_norm <- exclude_self_newclass_other %>% group_by(Cluster, animal, same_diff_type) %>% summarise(classes = sum(Freq))
same_type_norm$metric <- "pair_type"
same_type_norm <- same_type_norm %>% rename("same_diff" = "same_diff_type")
same_type_norm <- same_type_norm %>% merge(norm_to_other_class, on = "animal")


same_group_norm <- exclude_self_newclass_other %>% group_by(Cluster, animal, same_diff_group) %>% summarise(classes = sum(Freq))
same_group_norm$metric <- "group"
same_group_norm <- same_group_norm %>% rename("same_diff" = "same_diff_group")
same_group_norm <- same_group_norm %>% merge(norm_to_other_class, on = "animal")


same_parents_norm <- exclude_self_newclass_other %>% group_by(Cluster, animal, same_diff_parents) %>% summarise(classes = sum(Freq))
same_parents_norm$metric <- "parents"
same_parents_norm <- same_parents_norm %>% rename("same_diff" = "same_diff_parents")
same_parents_norm <- same_parents_norm %>% merge(norm_to_other_class, on = "animal")


# norm_to_other_class <- norm_to_other_class %>% merge(same_sex_norm, on = "animal") %>% merge(same_coh_norm, on = "animal") %>% merge(same_type_norm, on = "animal") %>% merge(same_group_norm, on = "animal") %>% merge(same_parents_norm, on = "animal")
norm_to_other_class <- rbind(same_sex_norm, same_coh_norm, same_type_norm, same_group_norm, same_parents_norm)
norm_to_other_class$pcts <- norm_to_other_class$classes/norm_to_other_class$sum_other_classes
```

normalize to the percent of classifications that are for "other" (i.e. what percent of "other" classifications?)
```{r}
norm_to_other_pct <- exclude_self_newclass_other %>% group_by(Cluster, animal) %>% summarise(sum_other_classes = sum(Freq))
exclude_self_newclass_other <- merge(exclude_self_newclass_other, norm_to_other_pct, on = "animal")
exclude_self_newclass_other$norm_to_other <- exclude_self_newclass_other$Freq/exclude_self_newclass_other$sum_other_classes
```

plot misclassification allocations by cluster
```{r, fig.width = 16}
clust_cols <- c("Drd1Pdyn" = "#304880",
                "Drd1PdynOprm1" = "#4898D0",
                "Drd1Penk" = "#88C0F0",
                "Drd2Penk" = "#9060A0",
                "Drd2NoPenk" = "#C078C0",
                "GABAergicNeurons" = "#389078",
                "Dlx2ImmatureNeurons" = "#70A830",
                "SstNpyInterneurons" = "#98D048",
                "PvalbInterneurons" = "#60D0A0",
                "CholinergicInterneurons" = "#80E8C0",
                "MatureOligos" = "#783028",
                "ImmatureOligos" = "#B82820",
                "Astrocytes" = "#D04058",
                "Microglia" = "#F87878")

clust_cols_new <- clust_cols
names(clust_cols_new) <- paste0(names(clust_cols), "_same")
clust_cols_light <- colorspace::lighten(clust_cols, 0.6)
clust_cols_light <- rep("gray69", 14)
names(clust_cols_light) <- paste0(names(clust_cols), "_different")

idx <- order(c(seq_along(clust_cols_new), seq_along(clust_cols_light)))
all_clust_cols <- (c(clust_cols_new,clust_cols_light))[idx]

newclass_long_pcts <- exclude_self_newclass_other %>% pivot_longer(cols = c("same_opp_sex", "same_diff_coh", "same_diff_type", "same_diff_group", "same_diff_parents"))

newclass_long_pcts$value <- factor(newclass_long_pcts$value, levels = c("same", "different"))
newclass_long_pcts$Cluster <- factor(newclass_long_pcts$Cluster, levels = f.levels)
newclass_long_pcts$Cluster_samediff <- paste0(newclass_long_pcts$Cluster, "_", newclass_long_pcts$value)
p <- ggplot(newclass_long_pcts, aes(x = factor(Cluster_samediff, levels = names(all_clust_cols)), y = norm_to_other*100, color = Cluster_samediff, fill = Cluster_samediff, alpha = 0.8)) +
  # geom_boxplot() +
  # geom_point(position = position_jitterdodge(jitter.width = 0, dodge.width = 0.9)) +
  geom_jitter(size = 0.5, width = 0.2) +
  scale_color_manual(values = all_clust_cols) +
  scale_fill_manual(values = all_clust_cols) +
  xlab("Group") +
  ylab("% of Classifications") +
  theme_classic()

p2 <- p + facet_wrap(~ name, nrow = 5)
print(p2)

# setwd("output/new_plots/")
# ggsave("by_metrics_animal_by_animal.pdf", p2)
```

plot allocations of misclassifications as barplot
```{r}
plt_list <- list()
for (m in c("same_diff_coh", "same_diff_group", "same_diff_type", "same_diff_parents", "same_opp_sex")) {
  print(m)
  #first find expected values
  metadata2 <- metadata
  if (m == "same_opp_sex") {
    metadata2 <- metadata2 %>% rename(m = "sex")
  }
  else if (m == "same_diff_group") {
    metadata2 <- metadata2 %>% rename(m = "Group")
  }
  else if(m == "same_diff_type") {
    metadata2 <- metadata2 %>% rename(m = "pair_type")
  }
  else if(m == "same_diff_parents") {
    metadata2 <- metadata2 %>% rename(m = "parents")
  }
  else if(m == "same_diff_coh") {
    metadata2 <- metadata2 %>% rename(m = "beh_cohort")
  }
  
  #get weighted avg expected value
  m_summ <- metadata2 %>% summarise(n_metric = n(), .by = m)
  m_summ$norm_pct <- m_summ$n_metric/38
  weighted_mean <- weighted.mean(m_summ$norm_pct, m_summ$n_metric)
  print(weighted_mean)
  
  #make df to sum all other classifications, don't group by cluster first
  test_m <- exclude_self_newclass_other %>% group_by_at(c("animal", "group", m)) %>% summarise_at(.vars = "Freq", .funs = sum)
  test_m <- test_m %>% rename("same_diff" = m)
  test_m2 <- test_m %>% group_by_at(c("animal", "group")) %>% summarise(total_other_classifications = sum(Freq)) %>% merge(test_m, on = "animal") %>% filter(same_diff == "same")
  test_m2$pct_same_classes <- test_m2$Freq/test_m2$total_other_classifications
  test_m2$metric <- m
  
  #one sample ttest vs. chance line
  one_samp_ttest <- t.test(test_m2$pct_same_classes*100, mu = weighted_mean*100)
  print(one_samp_ttest)
  
  p3 <- ggplot(test_m2, aes(x = metric, y = pct_same_classes*100)) + 
    geom_bar(stat = "summary", fun = "mean", color = "black", fill = "white") +
    stat_summary(
    fun.data = mean_se, fun.args = list(mult = 1), geom = "errorbar", color = "black", #fun.args = list(mult = 1), 
    position = position_dodge(0.8), width = 0.2, linewidth = 0.8
      ) +
    geom_jitter(height=0, width = 0.2, color = "black", alpha = 0.6, stroke = NA) +
    geom_hline(yintercept = weighted_mean*100, color = "red", linetype = "dashed") +
    ylim(0, 100) +
    theme_classic()
    print(p3)
  
  plt_list[[m]] <- p3
}

p4 <- grid.arrange(grobs = plt_list, nrow = 1)
print(p4)

setwd("output/new_plots/")
# ggsave("by_metrics_percent_cells_no_partner_combine_clusts_bars_se_noclustcollapsefirst.pdf", p4, width = 7, height = 7)
```

try a way to visualize only means (from data normalized to "other" counts)
this sets up dataframe to make a bubble plot
```{r}
norm_to_other_class2 <- norm_to_other_class %>% filter(same_diff == "same")

norm_for_summary <- norm_to_other_class2 %>% group_by(Cluster, metric) %>% summarise(mean_same_pct = mean(pcts, na.rm = TRUE))
```

bubble plot with normalized avg expected values
colored bubble indicates true mean
black line indicates expected (weighted mean)
```{r, fig.width = 8, fig.height = 6}
norm_wts <- data.frame()
for (m in unique(norm_for_summary$metric)) {
  print(m)
  m_df <- norm_for_summary %>% filter(metric == m)
  
  metadata2 <- metadata %>% filter(!animal %in% c("4918", "4967"))
  if (m == "cohort") {
    metadata2 <- metadata2 %>% rename(m = "beh_cohort")
  }
  
  else if (m == "group") {
    metadata2 <- metadata2 %>% rename(m = "Group")
  }
  
  #get weighted avg expected value
  m_summ <- metadata2 %>% summarise(n_metric = n(), .by = m)
  m_summ$norm_pct <- m_summ$n_metric/38
  weighted_mean <- weighted.mean(m_summ$norm_pct, m_summ$n_metric)
  
  mini_df <- data.frame("metric" = m, "weighted_mean" = weighted_mean)
  norm_wts <- rbind(norm_wts, mini_df)
}

norm_for_summary_weighted <- norm_for_summary %>% merge(norm_wts, on = "metric")
p <- ggplot(norm_for_summary_weighted, aes(x = metric, y = factor(Cluster, levels = rev(f.levels)), fill = Cluster, color = Cluster)) +
  geom_point(aes(size = mean_same_pct), alpha = 1, shape = 21) +
  geom_point(aes(size = weighted_mean), alpha = 1, shape = 1, color = "black") +
  # geom_point(aes(size = 1), alpha = 1, shape = 1, color = "black") + #if you want 100% circle
  scale_size_continuous(limits = c(0, 1), range = c(0, 10)) +
  scale_fill_manual(values = clust_cols, guide = FALSE) +
  scale_color_manual(values = clust_cols, guide = FALSE) +
  theme(panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = 1.2))

print(p)

# setwd("output/new_plots/")
# ggsave("SVM_metric_classifications_bubbleplot_normalized_expected.pdf", p)

```


```{r}
sessionInfo()
```





Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
