---
title: "UMAP Clustering and cell type proportions"
author: "Liza Brusman"
date: "2024-02-26"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r, message=FALSE}
#load packages
library(dplyr)
library(tidyr)
library(Seurat)
library(scCustomize)
library(ggplot2)
library(forcats)
library(ggpubr)
library(palettetown)
library(gridExtra)
library(glmmTMB)
library(DHARMa)
library(emmeans)
```

import previously saved seurat object
```{r}
#import seurat object. samples from all animals have already been integrated and non-nucleus accumbens cells have already been filtered out
#this file should be ~17.6 GB when imported
SCT_norm <- readRDS("output/SCT_norm.rds")
```

create aesthetic UMAP
```{r, fig.width = 9.5, fig.height = 7}
#pick new palette for UMAP
palette <- c("#304880", "#4898D0", "#88C0F0", "#9060A0", "#C078C0", "#389078", "#70A830", "#98D048", "#60D0A0", "#80E8C0", "#783028", "#B82820", "#D04058", "#F87878", "#F89040") 

#reorder clusters for plotting
order <- c("Drd1Pdyn", "Drd1PdynOprm1", "Drd1Penk", "Drd2Penk", "Drd2NoPenk", "GABAergicNeurons", "Dlx2ImmatureNeurons", "SstNpyInterneurons", "PvalbInterneurons", "CholinergicInterneurons", "MatureOligos", "ImmatureOligos", "Astrocytes", "Microglia", "RadialGlia-LikeCells")

DefaultAssay(SCT_norm) <- "SCT"
Idents(SCT_norm) <- "new_clusts"

# reorder clusters
Idents(SCT_norm) <- factor(Idents(SCT_norm), 
                            levels=order)

#create UMAP
ani.umap <- DimPlot(object = SCT_norm, reduction = "umap", label = FALSE, cols = palette, raster = FALSE)
print(ani.umap)

# setwd("output/")
# ggsave(
#   "umap_final.pdf",
#   plot = ani.umap,
#   device = pdf,
#   scale = 1,
#   width = 9.5,
#   height = 7,
#   units = c("in"),
#   dpi = 300,
#   limitsize = FALSE,
#   bg = "white"
# )

```

make dotplot of known marker genes
```{r, fig.width = 12, fig.height = 8}
# reorder clusters so they're in right order to plot
order <- rev(c("Drd1Pdyn", "Drd1PdynOprm1", "Drd1Penk", "Drd2Penk", "Drd2NoPenk", "GABAergicNeurons", "Dlx2ImmatureNeurons", "SstNpyInterneurons", "PvalbInterneurons", "CholinergicInterneurons", "MatureOligos", "ImmatureOligos", "Astrocytes", "Microglia", "RadialGlia-LikeCells"))

Idents(SCT_norm) <- "new_clusts"

Idents(SCT_norm) <- factor(Idents(SCT_norm), 
                            levels=order)

#set default assay to SCT
DefaultAssay(SCT_norm) <- "SCT"

markers.to.plot <- c("DRD1A", "DRD2", "Pdyn", "Penk", "Oprm1", "Igfbpl1", "Dlx2", "Sst", "Npy", "Vip", "Kit", "Chat", "Mog", "Olig2", "Pdgfra", "Gja1", "Aif1", "Vim", "Elavl2", "Gad1", "Syt1")

# #create dotplot with original colors
# dots <- DotPlot_scCustom(SCT_norm, features = markers.to.plot, dot.scale = 12, col.min=0, col.max=3, colors_use = viridis_plasma_light_high) + RotatedAxis()
# print(dots)

#create dotplot with new colors
dots <- DotPlot(SCT_norm, features = markers.to.plot, dot.scale = 12, col.min=0, col.max=3, cols = c("whitesmoke", "black")) + 
    RotatedAxis()
print(dots)

# setwd("output/")
# ggsave(
#   "dotplot_black.pdf",
#   plot = dots,
#   device = pdf,
#   scale = 1,
#   width = 12,
#   height = 8,
#   units = c("in"),
#   dpi = 300,
#   limitsize = TRUE,
#   bg = "white"
# )
```

get top marker genes for each cluster
```{r, eval=FALSE}
#find all markers
Idents(SCT_norm) <- "new_clusts"
markers <- FindAllMarkers(SCT_norm, only.pos=TRUE, min.pct=0.25, logfc.threshold=0.25)
markers <- markers %>% group_by(cluster) %>% arrange(desc(avg_log2FC), .by_group=TRUE)

# setwd("output/")
# write.csv(markers, "SCT_norm_markers.csv")
```

read in markers
```{r}
markers <- read.csv("output/SCT_norm_markers.csv")
```


make heatmap of top 10 marker genes for each cluster
this takes a long time to run
```{r, fig.width = 12, fig.height = 8}
DefaultAssay(SCT_norm) <- "SCT"
Idents(SCT_norm) <- "new_clusts"

order <- c("Drd1Pdyn", "Drd1PdynOprm1", "Drd1Penk", "Drd2Penk", "Drd2NoPenk", "GABAergicNeurons", "Dlx2ImmatureNeurons", "SstNpyInterneurons", "PvalbInterneurons", "CholinergicInterneurons", "MatureOligos", "ImmatureOligos", "Astrocytes", "Microglia", "RadialGlia-LikeCells")

palette <- c("#304880", "#4898D0", "#88C0F0", "#9060A0", "#C078C0", "#389078", "#70A830", "#98D048", "#60D0A0", "#80E8C0", "#783028", "#B82820", "#D04058", "#F87878", "#F89040") #"#98F088", 

# reorder clusters
Idents(SCT_norm) <- factor(Idents(SCT_norm),
                            levels=order)

#find top 10 markers for each cluster
markers %>% group_by(cluster) %>% top_n(n=10, wt=avg_log2FC) -> top10
top10 <- top10 %>% arrange(factor(cluster, levels = order))

heat <- DoHeatmap(SCT_norm, features = top10$gene, group.colors = palette) + NoLegend() + scale_fill_viridis_c(option = "magma") + theme(text = element_text(size = 8))

print(heat)

# setwd("output/")
# ggsave(
#   "heatmap.png",
#   plot = heat,
#   device = png,
#   scale = 1,
#   width = 12,
#   height = 8,
#   units = c("in"),
#   dpi = 300,
#   limitsize = TRUE,
#   bg = "white"
# )
```


stacked barplot of cell types in groups
```{r, fig.width = 8, fig.height = 10}
# reorder clusters
order <- c("Drd1Pdyn", "Drd1PdynOprm1", "Drd1Penk", "Drd2Penk", "Drd2NoPenk", "GABAergicNeurons", "Dlx2ImmatureNeurons", "SstNpyInterneurons", "PvalbInterneurons", "CholinergicInterneurons", "MatureOligos", "ImmatureOligos", "Astrocytes", "Microglia", "RadialGlia-LikeCells")
palette <- c("#304880", "#4898D0", "#88C0F0", "#9060A0", "#C078C0", "#389078", "#70A830", "#98D048", "#60D0A0", "#80E8C0", "#783028", "#B82820", "#D04058", "#F87878", "#F89040") #"#98F088",

Idents(SCT_norm) <- "new_clusts"
Idents(SCT_norm) <- factor(Idents(SCT_norm), 
                            levels=order)
#find proportions of each cell type per group
SCT_norm$Group <- paste(SCT_norm$Sex, SCT_norm$SSOS, sep = "_")
props <- prop.table(table(Idents(SCT_norm), SCT_norm$Group), margin = 2)
#x100 to get percentages
props <- props*100
#reorder columns to plot in correct order
props <- props[,c("F_SS", "F_OS", "M_SS", "M_OS")]

# Open a pdf file
# pdf("output/stackedbar.pdf", width = 8, height = 10)
#set font parameters for barplot
par(cex.axis=2, cex.lab=3)
# Get the stacked barplot
barplot(props, 
        col=palette ,
        border="white", 
        space=0.04, 
        font.axis=1,
        # cex.axis = 2,
        # cex.lab = 2,
        xlab="Group",
        ylab="Percent of Cells",
        las = 1
        # legend=rownames(props)
        )
#close pdf file
# dev.off()

```
import metadata
```{r}
metadata <- read.csv("../../docs/seq_beh_metadata.csv")

```


```{r}
#find proportions of each cell type per animal
Idents(SCT_norm) <- "new_clusts"
props_perani <- prop.table(table(Idents(SCT_norm), SCT_norm$Ani), margin = 2)
#x100 to get percentages
props_perani <- props_perani*100
#reorder columns to plot in correct order
# props <- props[,c("F_SS", "F_OS", "M_SS", "M_OS")]

props_perani <- props_perani %>% as.data.frame
props_perani <- props_perani %>% rename("animal" = "Var2")

props_perani <- merge(props_perani, metadata, on = "animal")
props_perani$Group <- paste(props_perani$sex, props_perani$SS_OS, sep = "_")

```

```{r, fig.height = 8, fig.width = 10}
# reorder clusters
order <- c("Drd1Pdyn", "Drd1PdynOprm1", "Drd1Penk", "Drd2Penk", "Drd2NoPenk", "GABAergicNeurons", "Dlx2ImmatureNeurons", "SstNpyInterneurons", "PvalbInterneurons", "CholinergicInterneurons", "MatureOligos", "ImmatureOligos", "Astrocytes", "Microglia", "RadialGlia-LikeCells")

setwd("output/")
#look at percentages by cluster
my_comparisons <- list( c("F_SS", "F_OS"), c("M_SS", "M_OS"), c("F_OS", "M_OS"), c("F_SS", "M_SS"))
for (clust in order) {
  print(clust)
  mini_df <- props_perani %>% filter(Var1 == clust)
  p <- mini_df %>% mutate(Group = fct_relevel(Group, "F_SS", "F_OS", "M_SS", "M_OS")) %>% ggplot(aes(x = Group, y = Freq, fill = Group, color = Group, alpha = 0.8)) + 
    geom_violin(lwd = 0.75) + 
    geom_point(position = position_dodge(width = 0.75), color = "slategrey", size = 2, alpha = 1) + 
    # stat_compare_means(comparisons= my_comparisons, paired = FALSE) +
    scale_fill_manual(values = c("F_SS" = "mediumpurple",
                                "F_OS"="darkorchid4",
                                "M_SS"="lightseagreen",
                                "M_OS"="deepskyblue4")) +
    scale_color_manual(values = c("F_SS" = "mediumpurple",
                                "F_OS"="darkorchid4",
                                "M_SS"="lightseagreen",
                                "M_OS"="deepskyblue4")) +
    ylim(0, NA) +
    ggtitle(clust) + 
    ylab("Percent of Cells") +
    xlab("Group") +
    theme_classic() +
    theme(text = element_text(size = 40))# + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
    print(p)

  
  # fname_pdf <- paste(clust, "props_violin.pdf", sep ="_")
  # fname_png <- paste(clust, "props_violin.png", sep ="_")
  # ggsave(fname_pdf, p, height = 8, width = 10, units = "in", device = pdf)
  # ggsave(fname_png, p, bg = "white", height = 8, width = 10, units = "in", device = png)
  }


```

faceted plots
```{r, fig.height = 8, fig.width = 10}
# reorder clusters
order <- c("Drd1Pdyn", "Drd1PdynOprm1", "Drd1Penk", "Drd2Penk", "Drd2NoPenk", "GABAergicNeurons", "Dlx2ImmatureNeurons", "SstNpyInterneurons", "PvalbInterneurons", "CholinergicInterneurons", "MatureOligos", "ImmatureOligos", "Astrocytes", "Microglia", "RadialGlia-LikeCells")

#look at percentages by cluster
my_comparisons <- list( c("F_SS", "F_OS"), c("M_SS", "M_OS"), c("F_OS", "M_OS"), c("F_SS", "M_SS"))
plt_list <- list()
for (clust in order) {
  print(clust)
  mini_df <- props_perani %>% filter(Var1 == clust)
  p <- mini_df %>% mutate(Group = fct_relevel(Group, "F_SS", "F_OS", "M_SS", "M_OS")) %>% ggplot(aes(x = Group, y = Freq, fill = Group, color = Group, alpha = 0.8)) + 
    geom_violin(lwd = 0.75) + 
    geom_point(position = position_dodge(width = 0.75), color = "slategrey", size = 1, alpha = 1) + 
    # stat_compare_means(comparisons= my_comparisons, paired = FALSE) +
    scale_fill_manual(values = c("F_SS" = "mediumpurple",
                                "F_OS"="darkorchid4",
                                "M_SS"="lightseagreen",
                                "M_OS"="deepskyblue4")) +
    scale_color_manual(values = c("F_SS" = "mediumpurple",
                                "F_OS"="darkorchid4",
                                "M_SS"="lightseagreen",
                                "M_OS"="deepskyblue4")) +
    ylim(0, NA) +
    ggtitle(clust) + 
    ylab("Percent of Cells") +
    xlab("Group") +
    theme_classic() +
    theme(text = element_text(size = 8))# + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
    print(p)

  plt_list[[clust]] <- p
}

all_plt <- grid.arrange(grobs = plt_list, nrow = 3)
all_plt

# setwd("output/")
# ggsave("all_cluster_props_violins.pdf", all_plt, width = 16, height = 8, units = "in")
```
do stats on violin plots
```{r, eval=FALSE}
props_perani_wide <- props_perani %>% pivot_wider(id_cols = c("animal", "sex", "SS_OS"), names_from = "Var1", values_from = "Freq")
all_contrasts <- data.frame(contrast = character(), 
                            estimate = numeric(), 
                            SE = numeric(), 
                            df = numeric(), 
                            t.ratio = numeric(), 
                            p.value = numeric(), 
                            Module = character())

for (clust in order) {
  form <- formula(paste0(clust, "~sex*SS_OS"))
  fit <- glmmTMB(form, data = props_perani_wide)
  print(summary(fit))
  
  # simres <- simulateResiduals(fit)
  # plot(simres, title = clust)
  
  EMM <- emmeans(fit, ~ sex*SS_OS)
  # print(summary(pairwise_comparisons))
  coef <- contrast(EMM, "pairwise")[c(1, 2, 5, 6)]
  coef2 <- summary(coef, adjust = "fdr")
  print(coef2)
  mini.df <- coef2 %>% as.data.frame()
  mini.df$Cluster <- clust
  
  all_contrasts <- rbind(all_contrasts, mini.df)
  
}
```

proportions of clusters within MSNs
```{r}
MSN_clusts <- c("Drd1Pdyn", "Drd1PdynOprm1", "Drd1Penk", "Drd2Penk", "Drd2NoPenk")
# MSN_clusts <- c("ImmatureOligos", "MatureOligos")


Idents(SCT_norm) <- "new_clusts"
SCT_MSN <- subset(SCT_norm, idents = MSN_clusts)

props_perani <- prop.table(table(Idents(SCT_MSN), SCT_MSN$Ani), margin = 2)
#x100 to get percentages
props_perani <- props_perani*100
#reorder columns to plot in correct order
# props <- props[,c("F_SS", "F_OS", "M_SS", "M_OS")]

props_perani <- props_perani %>% as.data.frame
props_perani <- props_perani %>% rename("animal" = "Var2")

props_perani <- merge(props_perani, metadata, on = "animal")
props_perani$Group <- paste(props_perani$sex, props_perani$SS_OS, sep = "_")


#plot
#look at percentages by cluster
# my_comparisons <- list( c("F_SS", "F_OS"), c("M_SS", "M_OS"), c("F_OS", "M_OS"), c("F_SS", "M_SS"))
plt_list <- list()
for (clust in MSN_clusts) {
  print(clust)
  mini_df <- props_perani %>% filter(Var1 == clust)
  p <- mini_df %>% mutate(Group = fct_relevel(Group, "F_SS", "F_OS", "M_SS", "M_OS")) %>% ggplot(aes(x = Group, y = Freq, fill = Group, color = Group, alpha = 0.8)) + 
    geom_violin(lwd = 0.75) + 
    geom_point(position = position_dodge(width = 0.75), color = "slategrey", size = 1, alpha = 1) + 
    # stat_compare_means(comparisons= my_comparisons, paired = FALSE) +
    scale_fill_manual(values = c("F_SS" = "mediumpurple",
                                "F_OS"="darkorchid4",
                                "M_SS"="lightseagreen",
                                "M_OS"="deepskyblue4")) +
    scale_color_manual(values = c("F_SS" = "mediumpurple",
                                "F_OS"="darkorchid4",
                                "M_SS"="lightseagreen",
                                "M_OS"="deepskyblue4")) +
    ylim(0, NA) +
    ggtitle(clust) + 
    ylab("Percent of Cells") +
    xlab("Group") +
    theme_classic() +
    theme(text = element_text(size = 8))# + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
    print(p)

  plt_list[[clust]] <- p
}

# #do stats
# props_perani_wide <- props_perani %>% pivot_wider(id_cols = c("animal", "sex", "SS_OS"), names_from = "Var1", values_from = "Freq")
# all_contrasts <- data.frame(contrast = character(), 
#                             estimate = numeric(), 
#                             SE = numeric(), 
#                             df = numeric(), 
#                             t.ratio = numeric(), 
#                             p.value = numeric(), 
#                             Module = character())
# 
# for (clust in MSN_clusts) {
#   form <- formula(paste0(clust, "~sex*SS_OS"))
#   fit <- glmmTMB(form, data = props_perani_wide)
#   print(summary(fit))
#   
#   # simres <- simulateResiduals(fit)
#   # plot(simres, title = clust)
#   
#   EMM <- emmeans(fit, ~ sex*SS_OS)
#   # print(summary(pairwise_comparisons))
#   coef <- contrast(EMM, "pairwise")[c(1, 2, 5, 6)]
#   coef2 <- summary(coef, adjust = "fdr")
#   print(coef2)
#   mini.df <- coef2 %>% as.data.frame()
#   mini.df$Cluster <- clust
#   
#   all_contrasts <- rbind(all_contrasts, mini.df)
#   
# }

```
proportion of all MSNs
```{r}
MSN_clusts <- c("Drd1Pdyn", "Drd1PdynOprm1", "Drd1Penk", "Drd2Penk", "Drd2NoPenk")

SCT_norm$MSN_or_no <- if_else(SCT_norm$new_clusts %in% MSN_clusts, "MSN", "not_MSN")

Idents(SCT_norm) <- "MSN_or_no"

props_perani <- prop.table(table(Idents(SCT_norm), SCT_norm$Ani), margin = 2)
#x100 to get percentages
props_perani <- props_perani*100
#reorder columns to plot in correct order
# props <- props[,c("F_SS", "F_OS", "M_SS", "M_OS")]

props_perani <- props_perani %>% as.data.frame
props_perani <- props_perani %>% rename("animal" = "Var2")

props_perani <- merge(props_perani, metadata, on = "animal")
props_perani$Group <- paste(props_perani$sex, props_perani$SS_OS, sep = "_")


mini_df <- props_perani %>% filter(Var1 == "MSN")
p <- mini_df %>% mutate(Group = fct_relevel(Group, "F_SS", "F_OS", "M_SS", "M_OS")) %>% ggplot(aes(x = Group, y = Freq, fill = Group, color = Group, alpha = 0.8)) + 
    geom_violin(lwd = 0.75) + 
    geom_point(position = position_dodge(width = 0.75), color = "slategrey", size = 1, alpha = 1) + 
    scale_fill_manual(values = c("F_SS" = "mediumpurple",
                                "F_OS"="darkorchid4",
                                "M_SS"="lightseagreen",
                                "M_OS"="deepskyblue4")) +
    scale_color_manual(values = c("F_SS" = "mediumpurple",
                                "F_OS"="darkorchid4",
                                "M_SS"="lightseagreen",
                                "M_OS"="deepskyblue4")) +
    ylim(0, NA) +
    ylab("Percent of Cells") +
    xlab("Group") +
    theme_classic() +
    theme(text = element_text(size = 8))# + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
print(p)

ggsave("output/all_msn_props.pdf", p)


# #do stats
# props_perani_wide <- props_perani %>% pivot_wider(id_cols = c("animal", "sex", "SS_OS"), names_from = "Var1", values_from = "Freq")
# all_contrasts <- data.frame(contrast = character(), 
#                             estimate = numeric(), 
#                             SE = numeric(), 
#                             df = numeric(), 
#                             t.ratio = numeric(), 
#                             p.value = numeric(), 
#                             Module = character())
# 
# form <- formula(paste0("MSN~sex*SS_OS"))
# fit <- glmmTMB(form, data = props_perani_wide)
# print(summary(fit))
# 
# EMM <- emmeans(fit, ~ sex*SS_OS)
# # print(summary(pairwise_comparisons))
# coef <- contrast(EMM, "pairwise")[c(1, 2, 5, 6)]
# coef2 <- summary(coef, adjust = "fdr")
# print(coef2)


```

sessioninfo
```{r}
sessionInfo()
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
