---
title: "Analysis of Hotspot data with behavior data"
author: "Liza Brusman"
date: "2024-10-31 👻"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r, message = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(Hmisc)
library(corrplot)
library(ggridges)
library(forcats)
library(stringr)
library(gprofiler2)
library(ComplexHeatmap)
library(lme4)
library(glmmTMB)
library(Matrix)
library(boot)
library(readxl)
library(emmeans)
```

```{r}
free_int_data <- readxl::read_excel("../../docs/free_int_beh.xlsx") %>% rename("free_int_time" = "Total Duration(Second)") %>% filter(pair != "4918x4967")
```

import metadata
```{r}
metadata <- read.csv("../../docs/seq_beh_metadata.csv")
metadata$Group <- paste(metadata$sex, metadata$SS_OS, sep = "_")

free_int_data <- free_int_data %>% filter(pair %in% metadata$pair) #%>% merge(metadata, on = "pair")
```

import ppt behavior
```{r}
all_beh_data <- read.csv("../../docs/PPTMetrics_coh1234_updated.csv")

all_beh_data2 <- all_beh_data %>% pivot_wider(id_cols=c("Animal.id", "Pair", "Treatment", "Sex", "Color"), names_from="Timepoint", values_from=c(16:24))

all_beh_data2 <- all_beh_data2 %>% rename("animal" = "Animal.id")
```

import module expression data from Hotspot
```{r}
lognorm_mods <- read.csv("../../docs/ani_mod_scores_allcells_lognorm_counts.csv") %>% merge(all_beh_data2, on = "animal") %>% filter(animal != "4967")

lognorm_meta <- lognorm_mods %>% merge(metadata, on = "animal")
```

get module names
```{r}
modules <- c()
for (i in 1:23) {
  mod_num <- paste0("Module.", i)
  modules <- modules %>% append(mod_num)
}
```

merge free interaction data, behavior data, and module expression
```{r}
free_int_data <- free_int_data %>% rename("Pair" = "pair")

lognorm_mods_beh <- lognorm_mods %>% merge(free_int_data[,c("Pair", "free_int_time")], on = "Pair") %>% merge(metadata, on = "animal")
```

make pairwise df
```{r}
pairwise_df <- data.frame(Pair = unique(lognorm_mods$Pair)) %>% filter(Pair != "4918x4967")
behs_mods <- lognorm_mods %>% select_if(is.numeric) %>% select(-c("animal")) %>% colnames()
for (i in behs_mods) {
  i_df <- lognorm_mods %>% pivot_wider(id_cols = Pair, names_from = Color, values_from = i)
  i_df[paste0(i, "_absdiff")] <- abs(i_df$O - i_df$B)
  i_df[paste0(i, "_avg")] <- (i_df$O + i_df$B)/2
  i_df <- i_df %>% select(Pair, paste0(i, "_absdiff"), paste0(i, "_avg"))
  pairwise_df <- merge(pairwise_df, i_df, on = "Pair")
}

pairs_types <- metadata[,c("pair", "pair_type")] %>% unique() %>% filter(pair != "4918x4967") %>% rename("Pair" = "pair")

pairwise_df <- pairwise_df %>% merge(pairs_types, on = "pair")

free_int_pairs <- free_int_data[,c("Pair", "free_int_time")] %>% merge(pairwise_df, on = "pair") # 

```

look at correlation between PPT data and module expression
```{r}
groups <- c("F_SS", "F_OS", "M_SS", "M_OS")

lognorm_mods_beh <- lognorm_mods_beh %>% merge(metadata[,c("animal", "pair_type", "SS_OS", "Group")], on = "animal")

for (group in groups) {
  group_df <- lognorm_mods_beh %>% filter(Group == group)

  to_plot <- group_df %>% select(contains("Module"), "huddle.time.partner_LT", "huddle.time.novel_LT")
  res <- rcorr(as.matrix(to_plot), type = "spearman")
  diag(res$P) <- 1
  
  behaviors <- c("huddle.time.partner_LT", "huddle.time.novel_LT") #, "percent.phuddle_LT" huddle.time.partner_LT", "huddle.time.novel_LT "free_int_time", 

  #for padj method
  Rhos_plot <- res$r[behaviors, modules] #%>% as.matrix()
  ps_plot <- res$P[behaviors, modules] #%>% as.matrix()

  # pdf(file = paste0("output/corrplot_seq_beh_individual_", type, ".pdf"), width = 12, height = 6)
  corrplot(Rhos_plot, 
           type = "full",
           method = "color",
           tl.col = "black", 
           addgrid.col = "white",
           col=colorRampPalette(c("#053061","white", "#67001F"))(200),
           tl.srt = 90,
           p.mat = ps_plot,
           sig.level=c(0.001, 0.01, 0.05),
           insig="label_sig",
           pch.cex=2,
           title=group)
  # dev.off()
}
```

DIY permutations/bootstrapping to find significant correlations
```{r, eval = FALSE}
set.seed(123)
groups <- c("F_SS", "F_OS", "M_SS", "M_OS")

#merge with metadata
lognorm_mods_beh <- lognorm_mods_beh %>% merge(metadata[,c("animal", "pair_type", "SS_OS", "Group")], on = "animal")

#set up df to hold summary stats
summary_stats <- data.frame()
for (group in groups) {
  print(group)
  #make separate df just for that group
  group_df <- lognorm_mods_beh %>% filter(Group == group)
  
  #pick out variables to plot in correlation matrix
  to_plot <- group_df %>% select(contains("Module"), "huddle.time.partner_LT", "huddle.time.novel_LT", "free_int_time")
  
  vars <- colnames(to_plot)
  
  #loop through all combinations of variables to run individual correlations
  for (v in vars) {
    for (w in vars) {
      c <- rcorr(to_plot[[v]], to_plot[[w]], type = "spearman")
      
      #then shuffle variables to determine the null distribution of correlations
      sub_shuffled_df <- data.frame()
      for (i in 1:1000) { #1:1000
        c2 <- rcorr(sample(to_plot[[v]]), sample(to_plot[[w]]), type = "spearman")
        
        #create mini df for that shuffle
        mini_shuffled_df <- data.frame("Var1" = v, "Var2" = w, "Rho_shuff" = c2$r[2], "pval_shuff" = c2$P[2], "n_iter" = i, "Group" = group)
        sub_shuffled_df <- rbind(sub_shuffled_df, mini_shuffled_df)
        # shuffled_df <- rbind(shuffled_df, mini_shuffled_df)
      }
      
      #create df with true values and means of shuffled values
      mini_df <- data.frame("Var1" = v, "Var2" = w, "Rho_true" = c$r[2], "pval_true" = c$P[2], 
                            "Rho_shuff" = mean(sub_shuffled_df$Rho_shuff), "pval_shuff" = mean(sub_shuffled_df$pval_shuff))
      
      #find top 1% of p-values
      n <- 1
      pval_top1perc_df <- sub_shuffled_df[sub_shuffled_df$pval_shuff < quantile(sub_shuffled_df$pval_shuff, prob=0.01),]
      pval_top1perc <- max(pval_top1perc_df$pval_shuff)
      #find top 5% of p-values
      n <- 5
      pval_top5perc_df <- sub_shuffled_df[sub_shuffled_df$pval_shuff < quantile(sub_shuffled_df$pval_shuff, prob=0.05),]
      pval_top5perc <- max(pval_top5perc_df$pval_shuff)
      
      #create df with all these summary stats
      summ_df <- data.frame(Group = group,
                           Var1 = v,
                           Var2 = w,
                           n_samps = nrow(sub_shuffled_df),
                           Rho_shuff_mean = mean(sub_shuffled_df$Rho_shuff),
                           Rho_shuff_stdev = sd(sub_shuffled_df$Rho_shuff),
                           Rho_shuff_sterr = sd(sub_shuffled_df$Rho_shuff)/sqrt(nrow(sub_shuffled_df)),
                           pval_shuff_mean = mean(sub_shuffled_df$pval_shuff),
                           pval_shuff_stdev = sd(sub_shuffled_df$pval_shuff),
                           pval_shuff_sterr = sd(sub_shuffled_df$pval_shuff)/sqrt(nrow(sub_shuffled_df)),
                           Rho_true = c$r[2],
                           pval_true = c$P[2],
                           pval_cutoff_1perc = pval_top1perc,
                           pval_cutoff_5perc = pval_top5perc
                           )
      summary_stats <- rbind(summary_stats, summ_df)
    }
  }
}

```

save p-values as csv
```{r, eval = FALSE}
write.csv(summary_stats, "output/shuffled_corrs_behs_mods.csv")
```

read in csv
```{r}
summary_stats <- read.csv("output/shuffled_corrs_behs_mods.csv")
```


cutoff p-values to be within the top (lowest) 5% of shuffled p-values. these are the p-values plotted on the heatmaps
```{r}
sig_corrs <- summary_stats %>% filter(pval_true < pval_cutoff_5perc) %>% filter(Var1 == "huddle.time.partner_LT" | Var1 == "huddle.time.novel_LT") %>% filter(Var1 != Var2)
```


does free interaction time differ by pairing type?
```{r}
#make violin plot
p <- ggplot(free_int_pairs, aes(x = pair_type, y = free_int_time/60, color = pair_type, fill = pair_type)) + 
  geom_violin(alpha = 0.8, lwd = 1) + 
  geom_point(color = "slategrey", alpha = 1) + 
  scale_color_manual(values = c("FM" = "coral", "FF" = "slateblue", "MM" = "#1B9E77")) +
  scale_fill_manual(values = c("FM" = "coral", "FF" = "slateblue", "MM" = "#1B9E77")) +
  ylim(0, 7000/60) +
  xlab("Group") +
  ylab("Free interaction time (min)") +
  theme_classic()
print(p)
# ggsave("output/group_diffs_free_int.pdf", p)

#stats to see if there are group differences
fit <- glmmTMB(free_int_time~pair_type, data = free_int_pairs)
print(summary(fit))

EMM <- emmeans(fit, ~ pair_type)
coef <- contrast(EMM, "pairwise")
coef2 <- summary(coef, adjust = "fdr") #[["sex*SSOS"]])
print(coef2)
```




look at correlation between free interaction time and module expression
```{r}
for (mod in modules) {
  c <- rcorr(lognorm_mods_beh[,mod], lognorm_mods_beh$free_int_time/60, type = "spearman")
  # print(c)

  p <- ggplot(lognorm_mods_beh, aes(x = .data[[mod]], y = free_int_time/60)) + 
    geom_point(aes(color = pair_type)) + 
    geom_smooth(method = "lm", color = "black")  +
    scale_color_manual(values = c("FM" = "coral", "FF" = "slateblue", "MM" = "#1B9E77")) +
    ggtitle(paste0("Rho = ", round(c$r[2], 4), " p = ", round(c$P[2], 4))) +
    xlab(paste0(mod, " Expression")) +
    ylab("Free interaction time (min)") +
    theme_classic() +
    theme(aspect.ratio = 1) +
    theme(text = element_text(size = 16))
  
  print(p)
  
  # fname <- paste0("output/module", mod, "_free_int_corr.pdf")
  # ggsave(fname, p)
}


```

look at correlation between PPT and free interaction behavior
```{r}
#spearman correlation
c <- rcorr(lognorm_mods_beh$LT_phuddle, lognorm_mods_beh$free_int_time, type = "spearman")

#scatter plot
p <- ggplot(lognorm_mods_beh, aes(x = LT_phuddle/60, y = free_int_time/60)) + 
    geom_point(aes(color = pair_type)) + 
    geom_smooth(method = "lm", color = "black")  +
    scale_color_manual(values = c("FM" = "coral", "FF" = "slateblue", "MM" = "#1B9E77")) +
    ggtitle(paste0("Rho = ", round(c$r[2], 4), " p = ", round(c$P[2], 4))) +
    xlab("Partner Huddle Time (min)") +
    ylab("Free interaction time (min)") +
    theme_classic() +
    theme(aspect.ratio = 1) +
    theme(text = element_text(size = 16))
print(p)

#save
# ggsave("output/ppt_vs_freeint.pdf", p)

```


```{r}
sessionInfo()
```

