---
title: "Hotspot for paper"
author: "Liza Brusman"
date: "2024-10-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```


```{r, message=FALSE}
library(dplyr)
library(tidyr)
library(svMisc)
library(cowplot)
library(ggplot2)
library(pheatmap)
library(reshape2)
library(gridExtra)
library(RColorBrewer)
library(ivmte)
library(viridis)
library(palettetown)
library(ComplexHeatmap)
library(corrplot)
library(Hmisc)
library(corrr)
library(ggcorrplot)
library(FactoMineR)
library(stringr)
library(ggpubr)
library(ComplexHeatmap)
library(glmmTMB)
library(DHARMa)
library(emmeans)
library(Seurat)
source("hotspot.R")
```

import module data and metadata
```{r}
all_cells <- read.csv("../../docs/ani_mod_scores_allcells_lognorm_counts.csv") %>% filter(animal != "4967")

metadata <- read.csv("../../docs/seq_beh_metadata.csv") %>% process_metadata() %>% group_level()
```
merge meta and module data
```{r}
merged_data <- merge(all_cells, metadata, on = "animal")
```

```{r}
modules <- get_modules(merged_data)

modules <- c()
for (i in 1:23) {
  mod_num <- paste0("Module.", i)
  modules <- modules %>% append(mod_num)
}

```
get which genes are in which module
```{r}
mod_genes <- read.csv("../../docs/new_clusts_hotspot-gene-modules.csv")

```

import seurat object
```{r}
SCT_norm <- readRDS("../seurat_clustering/output/SCT_norm.rds")
```

outlier test for modules 5 and 12 (and Module 14 to check for new significance)
https://statsandr.com/blog/outliers-detection-in-r/
```{r}
ggplot(merged_data, aes(x = Group, y = Module.5)) + geom_boxplot()
out <- boxplot.stats(merged_data$Module.5)$out
out_ind <- which(merged_data$Module.5 %in% c(out))
out_ind

#try outlier test using percentiles
lower_bound <- quantile(merged_data$Module.5, 0.025)
lower_bound
upper_bound <- quantile(merged_data$Module.5, 0.975)
upper_bound

#for just SS females
SS_F <- merged_data %>% filter(Group == "F_SS")
ggplot(SS_F, aes(x = Group, y = Module.5)) + geom_boxplot()
out <- boxplot.stats(SS_F$Module.5)$out
out_ind <- which(SS_F$Module.5 %in% c(out))
out_ind

#try outlier test using percentiles
lower_bound <- quantile(SS_F$Module.5, 0.025)
lower_bound
upper_bound <- quantile(SS_F$Module.5, 0.975)
upper_bound

print("Module.5")
for_mod5 <- merged_data %>% filter(Module.5 != out)

my_comparisons <- list(c("F_SS", "F_OS"),
                         c("F_SS", "M_SS"),
                         c("M_SS", "M_OS"),
                         c("F_OS", "M_OS"))
plt <- ggplot(for_mod5, aes_string(x = "Group", y = "Module.5", color = "Group",
                                 fill = "Group", alpha = 0.8)) +
    geom_violin(lwd = 0.5) +
    geom_point(position = position_dodge(width = 0.75),
               color = "slategrey", size = 2, alpha = 1) +
    # stat_compare_means(comparisons = my_comparisons,
    #                    paired = FALSE, method = "wilcox.test") +

    scale_fill_manual(values = c("F_SS" = "mediumpurple",
                                 "F_OS" = "darkorchid4",
                                 "M_SS" = "lightseagreen",
                                 "M_OS" = "deepskyblue4")) +
    scale_color_manual(values = c("F_SS" = "mediumpurple",
                                  "F_OS" = "darkorchid4",
                                  "M_SS" = "lightseagreen",
                                  "M_OS" = "deepskyblue4")) +

    ylab("Module Score") +
    ggtitle("Module.5") +
    theme_classic() +
    theme(text = element_text(size = 40))
  print(plt)
  
fit <- glmmTMB(Module.5~Group, data = for_mod5)
print(summary(fit))
  
  # simres <- simulateResiduals(fit)
  # plot(simres, title = mod)

EMM <- emmeans(fit, ~ Group)
coef <- contrast(EMM, "pairwise")[c(1, 2, 5, 6)]
coef2 <- summary(coef, adjust = "fdr") #[["sex*SSOS"]])
print(coef2)


##for moudle.12
print("Module.12")

#for just SS females
SS_F <- merged_data %>% filter(Group == "F_SS")
ggplot(SS_F, aes(x = Group, y = Module.12)) + geom_boxplot()
out <- boxplot.stats(SS_F$Module.12)$out
out_ind <- which(SS_F$Module.12 %in% c(out))
out_ind

#try outlier test using percentiles
lower_bound <- quantile(SS_F$Module.12, 0.025)
lower_bound
upper_bound <- quantile(SS_F$Module.12, 0.975)
upper_bound

for_mod12 <- merged_data %>% filter(Module.12 != out)

my_comparisons <- list(c("F_SS", "F_OS"),
                         c("F_SS", "M_SS"),
                         c("M_SS", "M_OS"),
                         c("F_OS", "M_OS"))
plt <- ggplot(for_mod12, aes_string(x = "Group", y = "Module.12", color = "Group",
                                 fill = "Group", alpha = 0.8)) +
    geom_violin(lwd = 0.5) +
    geom_point(position = position_dodge(width = 0.75),
               color = "slategrey", size = 2, alpha = 1) +
    # stat_compare_means(comparisons = my_comparisons,
                       # paired = FALSE, method = "wilcox.test") +

    scale_fill_manual(values = c("F_SS" = "mediumpurple",
                                 "F_OS" = "darkorchid4",
                                 "M_SS" = "lightseagreen",
                                 "M_OS" = "deepskyblue4")) +
    scale_color_manual(values = c("F_SS" = "mediumpurple",
                                  "F_OS" = "darkorchid4",
                                  "M_SS" = "lightseagreen",
                                  "M_OS" = "deepskyblue4")) +

    ylab("Module Score") +
    ggtitle("Module.12") +
    theme_classic() +
    theme(text = element_text(size = 40))
  print(plt)
  
fit <- glmmTMB(Module.12~Group, data = for_mod12)
print(summary(fit))
  
  # simres <- simulateResiduals(fit)
  # plot(simres, title = mod)

EMM <- emmeans(fit, ~ Group)
coef <- contrast(EMM, "pairwise")[c(1, 2, 5, 6)]
coef2 <- summary(coef, adjust = "fdr") #[["sex*SSOS"]])
print(coef2)


##just look for Module-14
print("Module.14")

#for just SS females
SS_F <- merged_data %>% filter(Group == "F_SS")
ggplot(SS_F, aes(x = Group, y = Module.14)) + geom_boxplot()
out <- boxplot.stats(SS_F$Module.14)$out
out_ind <- which(SS_F$Module.14 %in% c(out))
out_ind

#try outlier test using percentiles
lower_bound <- quantile(SS_F$Module.14, 0.025)
lower_bound
upper_bound <- quantile(SS_F$Module.14, 0.975)
upper_bound

for_mod14 <- merged_data %>% filter(!Module.14 %in% out)

my_comparisons <- list(c("F_SS", "F_OS"),
                         c("F_SS", "M_SS"),
                         c("M_SS", "M_OS"),
                         c("F_OS", "M_OS"))
plt <- ggplot(for_mod14, aes_string(x = "Group", y = "Module.14", color = "Group",
                                 fill = "Group", alpha = 0.8)) +
    geom_violin(lwd = 0.5) +
    geom_point(position = position_dodge(width = 0.75),
               color = "slategrey", size = 2, alpha = 1) +
    # stat_compare_means(comparisons = my_comparisons,
                       # paired = FALSE, method = "wilcox.test") +

    scale_fill_manual(values = c("F_SS" = "mediumpurple",
                                 "F_OS" = "darkorchid4",
                                 "M_SS" = "lightseagreen",
                                 "M_OS" = "deepskyblue4")) +
    scale_color_manual(values = c("F_SS" = "mediumpurple",
                                  "F_OS" = "darkorchid4",
                                  "M_SS" = "lightseagreen",
                                  "M_OS" = "deepskyblue4")) +

    ylab("Module Score") +
    ggtitle("Module.14") +
    theme_classic() +
    theme(text = element_text(size = 40))
  print(plt)
  
fit <- glmmTMB(Module.14~Group, data = for_mod14)
print(summary(fit))
  
  # simres <- simulateResiduals(fit)
  # plot(simres, title = mod)

EMM <- emmeans(fit, ~ Group)
coef <- contrast(EMM, "pairwise")[c(1, 2, 5, 6)]
coef2 <- summary(coef, adjust = "fdr") #[["sex*SSOS"]])
print(coef2)
```


```{r}
for (mod in modules) {
  violin(merged_data, save = FALSE)
}
```

correlation plots
```{r}
setwd("output/")
for (mod in "Module.6") {
  corr_plot(merged_data, color = "deeppink4", font_size = 20, save = FALSE)
}

for (mod in "Module.11") {
  corr_plot(merged_data, color = "chartreuse4", font_size = 20, save = FALSE)
}
```
facet corr plots for easier manipulation in illustrator
```{r}
plt_list <- list()
for (mod in modules) {
  p <- corr_plot_bypair(merged_data, font_size = 8)
  plt_list[[mod]] <- p
}


all_plt <- grid.arrange(grobs = plt_list, nrow = 6)

# setwd("output/")
# ggsave("all_module_corrs.pdf", all_plt, width = 16, height = 16, units = "in")
```

make a plot of all between-partner rank distances separated by pair type
```{r}
all_dists_df <- make_dist_df(pairs, type, fake_real, dist, Module)
plt_list <- list()
for (mod in modules) {
    mod_dist_df <- get_rank_distances(merged_data)
    all_dists_df <- rbind(all_dists_df, mod_dist_df)
    
    p <- plot_rank_group(mod_dist_df)
    plt_list[[mod]] <- p
  }

plot_rank(all_dists_df)

all_plt <- grid.arrange(grobs = plt_list, nrow = 6)

# setwd("output/")
# ggsave("all_module_dists.pdf", all_plt, width = 18, height = 18, units = "in")

```

euclidean distances between partners and all non-partner pairs
```{r}
fake_pairs <- get_fake_pairs(merged_data)

all_euclidean_dists <- data.frame()
for (mod in modules) {
  mod_euclidean_dists <- get_euclidean_dists(SCT_norm, mod_genes, metadata)
  all_euclidean_dists <- rbind(all_euclidean_dists, mod_euclidean_dists)
}
```
plot euclidean distances
```{r}
setwd("output/")

plt_list <- list()
for (mod in modules) {
  p <- plot_euclidean_dists(all_euclidean_dists, save = FALSE)
  plt_list[[mod]] <- p

}

all_plt <- grid.arrange(grobs = plt_list, nrow = 6)

# ggsave("all_module_euclidean_dists.pdf", all_plt, width = 18, height = 18, units = "in")
```
plot euclidean distances by pair type
```{r}
plt_list <- list()
for (mod in modules) {
 p <- plot_euclidean_dists_pairtype(all_euclidean_dists)
 plt_list[[mod]] <- p
}

all_plt <- grid.arrange(grobs = plt_list, nrow = 6)

# setwd("output/")
# ggsave("all_module_euclidean_dists_by_pairtype.pdf", all_plt, width = 18, height = 18, units = "in")
```

summary stats from hotspot mods
```{r}
summs <- mod_genes %>% group_by(Module) %>% summarise(n_genes = n())
```

GLM for module expression by sex and SS_OS
```{r}
all_contrasts <- data.frame(contrast = character(), 
                            estimate = numeric(), 
                            SE = numeric(), 
                            df = numeric(), 
                            t.ratio = numeric(), 
                            p.value = numeric(), 
                            Module = character())

for (mod in modules) {
  form <- formula(paste0(mod, "~sex*SS_OS"))
  fit <- glmmTMB(form, data = merged_data)
  print(summary(fit))
  
  ## if you want to simulate and plot residuals
  # simres <- simulateResiduals(fit)
  # plot(simres, title = mod)
  
  EMM <- emmeans(fit, ~ sex*SS_OS)
  coef <- contrast(EMM, "pairwise")[c(1, 2, 5, 6)]
  coef2 <- summary(coef, adjust = "fdr")
  print(coef2)
  mini.df <- coef2 %>% as.data.frame()
  mini.df$Module <- mod
  
  all_contrasts <- rbind(all_contrasts, mini.df)
  
}
```
```{r}
# setwd("output/")
# write.csv(all_contrasts, "all_glm_posthocs_module_expr.csv")
```
stats for rank distances between groups
```{r}
all_contrasts_dists <- data.frame(contrast = character(), 
                            estimate = numeric(), 
                            SE = numeric(), 
                            df = numeric(), 
                            t.ratio = numeric(), 
                            p.value = numeric(), 
                            Module = character())

for (mod in modules) {
  mod_df <- all_dists_df %>% filter(Module == mod)
  fit <- glmmTMB(dist~type, data = mod_df)
  print(summary(fit))
  
  # simres <- simulateResiduals(fit)
  # plot(simres, title = mod)

  EMM <- emmeans(fit, ~ type)
  coef <- contrast(EMM, "pairwise")
  coef2 <- summary(coef, adjust = "fdr") #[["sex*SSOS"]])
  print(coef2)
  mini.df <- coef2 %>% as.data.frame()
  mini.df$Module <- mod
  
  all_contrasts_dists <- rbind(all_contrasts_dists, mini.df)
  
}
```
```{r}
# setwd("output/")
# write.csv(all_contrasts_dists, "glm_posthocs_rank_group_comparisons.csv")
```

wilcox.test for euclidean dists
```{r}
all_euclidean_stats <- data.frame()
for (mod in modules) {
  mod_test <- all_euclidean_dists %>% filter(Module == mod)
  mod_wilcox <- wilcox.test(dist ~ fake_real, data = mod_test)
  wilcox_df <- data.frame(Module = mod,
                          W = mod_wilcox$statistic[1],
                          p.value = mod_wilcox$p.value,
                          test = "wilcoxon test")
  all_euclidean_stats <- rbind(all_euclidean_stats, wilcox_df)
}
```
```{r}
# setwd("output/")
# write.csv(all_euclidean_stats, "euclidean_dists_fakereal_wilcoxtest.csv")
```


all module correlations stats
```{r}
corr_stats_df <- data.frame()
for (mod in modules) {
  mod_df <- merged_data %>% pivot_wider(id_cols = "pair", names_from = "color", values_from = mod)
  mod_cor <- rcorr(mod_df$O, mod_df$B, type = "spearman")
  mod_corr_stats <- data.frame(Module = mod,
                               Rho = mod_cor$r[2],
                               p.value = mod_cor$P[2],
                               test = "Spearman Correlation")
  corr_stats_df <- rbind(corr_stats_df, mod_corr_stats)
}
```
```{r}
# setwd("output/")
# write.csv(corr_stats_df, "module_partner_corrs_stats.csv")
```

```{r}
sessionInfo()
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
